### 2015-12-02 ch02-qlogms/excitation-design.tex
Example: numerical precision -> period

  \begin{example}\label{eg:excitation:non-periodic}
   Consider the generalized multisine of the previous example to which a spectral component is added at $f_3 = \pi \unit{Hz}$, or equivalently one with a period of $T_3 = \pi^{-1} \unit{s}$.
   Then it is not possible to determine integer values $n_1$, $n_2$, $n_3$ such that $n_1 T_1 = n_2 T_2 = n_3 T_3$ since $T_3$ is an irrational number of seconds (unlike $T_1$ and $T_2$).
   Consequently, such a generalized `multisine' is not periodic.
  \end{example}
  \begin{remark}
  In practical applications each frequency will be represented as either a fixed-point or floating point number with a limited numerical precision $\delta$.
  Hence, each frequency line $f_k$ will be approximated by a line at $\round{f_k \delta^{-1}} \delta \isdef m_k \delta$ where $m_k \in \Integers$.
  The resulting multisine has a period 
  \begin{equation}
     T 
     = \lcm \Set{T_k |  k \in \mathbb{K}} 
     = \lcm \Set{m_k^{-1} \delta^{-1} |  k \in \mathbb{K}}
     = \delta^{-1} \lcm \Set{m_k^{-1} |  k \in \mathbb{K}}
     \text{.}
  \end{equation}
  The latter $\lcm$ always exists because $m_k^{-1}  \in \RationalNumbers$ as $m_k \in \Integers$.
  The resulting period $T$ might, however, might be too long for practical use.
  E.g. when $\delta \approx 2.2\cdot 10^{-16}$ (\code{double} precision), the period of a multisine with components as in Example~\ref{eg:excitation:non-periodic} is $T \approx 2.8 \cdot 10^{14} \unit{s} \approx 9 \cdot 10^{6} \unit{year}$.
  \end{remark}

@@

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} or the \code{logspace} function in \MATLAB it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$).
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The~\citep{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

>>>>

  \begin{example}\label{eg:excitation:non-periodic}
   Consider the generalized multisine of the previous example to which a spectral component is added at $f_3 = \pi \unit{Hz}$, or equivalently one with a period of $T_3 = \pi^{-1} \unit{s}$.
   Then it is not possible to determine integer values $n_1$, $n_2$, $n_3$ such that $n_1 T_1 = n_2 T_2 = n_3 T_3$ since $T_3$ is an irrational number of seconds (unlike $T_1$ and $T_2$).
   Consequently, such a generalized `multisine' is not periodic.
  \end{example}
  \begin{remark}
  In practical applications each frequency will be represented as either a fixed-point or floating point number with a limited numerical precision $\eps$.
  Hence, each frequency line $f_k$ will be approximated by a line at $\round{f_k \eps^{-1}} \eps \isdef m_k \eps$ where $m_k \in \Integers$.
  The resulting multisine has a period 
  \begin{equation}
     T 
     = \lcm \Set{T_k |  k \in \mathbb{K}} 
     = \lcm \Set{m_k^{-1} \eps^{-1} |  k \in \mathbb{K}}
     = \eps^{-1} \lcm \Set{m_k^{-1} |  k \in \mathbb{K}}
     \text{.}
  \end{equation}
  The latter $\lcm$ always exists because $m_k^{-1}  \in \RationalNumbers$ as $m_k \in \Integers$.
  The resulting period $T$ might, however, might be too long for practical use.
  E.g. when $\eps \approx 1.2\cdot 10^{-7}$ (\code{single} precision), the period of a multisine with components as in Example~\ref{eg:excitation:non-periodic} is $T \approx 8.4 \cdot 10^{6} \unit{s} \approx 97 \unit{days}$.
  For \code{double} precision, $\eps \approx 2.2 \cdot 10^{-16}$ and the period $T \approx 9 \cdot 10^{6} \unit{year}$.
  \end{remark}

@@

  \begin{property}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} or the \code{logspace} function in \MATLAB it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$).
    Obviously, such alternative specifications are equivalent.
  \end{remark}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}
